Clustering and Classification of Satellite Imagery
----------------------------------------------------
# Unsupervised Clustering
This section will use unsupervised clustering to split up given satellite images into several groups,
with the goal of identifying meaningfull clusters that represent different geographic features such as water, farm land, forest or buildings.

```{r, echo=FALSE}
# Set up
#setwd("C:\\Users\\D059331\\Desktop\\DM GIC\\data\\img")
require('raster')
require(sp)
require(rgdal)
require(randomForest)
# Normalize a raster dataset

raster.scale <- function(rst, norm="min-max") {
	require(raster)
	nbands <- nlayers(rst)
	if (norm == "min-max") {
		Min <- vector(mode="numeric", length=nbands)
		Max <- vector(mode="numeric", length=nbands)
		for (k in 1:nbands) {
			Min[k] <- rst[[k]]@data@min
			Max[k] <- rst[[k]]@data@max
		}
		return(scale(rst, center=Min, scale=Max-Min))
	}
	# by default mean-std normalization
	return(scale(rst))
}


# Compute normalized difference (e.g. NDVI)
normdiff <- function(stk, red=3, nir=4, stack.it=FALSE, normalize=FALSE, norm="min-max") {
	require(raster)
	dims <- dim(stk)
	max.bands = max(red,nir)
	if (dims[3] < max.bands)
		stop("Raster file does not have enough bands\n")
	nirBand <- getValues(stk[[nir]])
	redBand <- getValues(stk[[red]])
	ndviBand <- (nirBand - redBand) / (nirBand + redBand)
	if (normalize == TRUE) {
		if (norm == "min-max") {
			center <- min(ndviBand)
			ampl <- max(ndviBand)-center
		} else {
			center <- mean(ndviBand)
			ampl <- sd(ndviBand)
		}
		ndviBand <- scale(ndviBand, center=center, scale=ampl)
	}
	ndvi <- raster(nrows=dims[1], ncols=dims[2], crs=crs(stk), ext=extent(stk), resolution=res(stk), vals=ndviBand)
	if (stack.it == FALSE)
		return(ndvi)
	return(stack(stk, ndvi))
}

# Moving function
# This function computes the moving (average, standard-deviation, etc) of a raster dataset
mov.fun <- function(rst, window.size, fun, normalize=FALSE, verbose=FALSE) {
	require(raster)
	if (!is.integer(window.size))
		window.size <- abs(ceiling(window.size))
	# need to be odd
	if (window.size%%2 == 0)
		window.size <- window.size + 1
	w <- matrix(1, nrow=window.size, ncol=window.size)
	n <- nlayers(rst)
	stk <- focal(rst[[1]], w, fun, na.rm=TRUE)
	if (verbose == TRUE)
		cat(">> Band ", 1," of ", n, " done.\n")
	for (k in 2:n) {
		tmp <- focal(rst[[k]], w, fun, na.rm=TRUE)
		stk <- stack(stk, tmp)
		if (verbose == TRUE)
			cat(">> Band ", k, " of ", n, " done.\n")
	}
	if (normalize == TRUE)
		stk <- raster.scale(stk)
	return(stk)
}

# function to calculate and plot kmeans for a raster object
performKMeans <- function(inputRaster, noClusters) {
	raster_df <- as.data.frame(inputRaster)
	clustering <- kmeans(raster_df, noClusters, iter.max = 100, nstart = 10)
	clusterRaster <- raster(inputRaster)
	clusterRaster <- setValues(clusterRaster, clustering$cluster)
	plot(clusterRaster)
	#return(clustering)
}

calculateError <- function(prediction, actual) {
	trainDiff <- predictionMatch - train[,7]
	trainDiffCount <- 0
	for(i in 1:length(trainDiff)) {
		if(trainDiff[i] != 0) {
			trainDiffCount <- trainDiffCount + 1
		}
	}
}
```

- Load a raster file, which has 4 layers
```{r}
rasterJ<-brick("C:\\Users\\D059331\\Desktop\\DM GIC\\data\\img\\J_04AUG14112729-M2AS-000000137917_01_P001_etrs89.TIF")
rasterJ
```
- Plot map
```{r}
plotRGB(rasterJ, 3,2,1)
```

- choose a sector for clustering
```{r}
ext <- extent(-104294.4, -102964.5, -43623.48, -42742.44 )
sector <- crop(rasterJ, ext)
plotRGB(sector, 3, 2, 1)
```

- add normdiff layer for better vegetation recognition
```{r}
sector[[5]] <- normdiff(sector)
```
- plot each layer separately
```{r}
plot(sector)
```

- check for NAs
```{r}
summary(sector)
```
- use only nir and ndvi layer for k-means
```{r}
sector_mod <- brick(sector[[4]],sector[[5]])
```

- perform kmeans and plot result
```{r}
performKMeans(sector_mod, 12)
```

K-Means doesnt seem to perform very well on this sector. Ocean and land arent clearly separated. Maybe this works better with fewer clusters (i.e. lower k).
```{r}
performKMeans(sector_mod, 6)
```

Lets try instead a different sector, one without ocean.
```{r}
ext <- extent(-101751, -100703.7, -47812.61, -46914.94)
landSector <- crop(rasterJ, ext)
landSector[[5]] <- normdiff(landSector)
plotRGB(landSector, 3, 2, 1)
landSector_mod <- brick(landSector[[4]],landSector[[5]])
performKMeans(landSector_mod, 12)
```

This is not too awesome. Try 6 clusters.
```{r}
performKMeans(landSector_mod, 12)
```