Clustering and Classification of Satellite Imagery
----------------------------------------------------
#Abstract
-RAQUEL
# Unsupervised Clustering
-RAQUEL any intro stuff about this topic, feell free to use references
This section will use unsupervised clustering to split up given satellite images into several groups,
with the goal of identifying meaningfull clusters that represent different geographic features such as water, farm land, forest or buildings.

```{r, echo=FALSE}
# Set up
#setwd("C:\\Users\\D059331\\Desktop\\DM GIC\\data\\img")
require('raster')
require(sp)
require(rgdal)
require(randomForest)
# Normalize a raster dataset

raster.scale <- function(rst, norm="min-max") {
	require(raster)
	nbands <- nlayers(rst)
	if (norm == "min-max") {
		Min <- vector(mode="numeric", length=nbands)
		Max <- vector(mode="numeric", length=nbands)
		for (k in 1:nbands) {
			Min[k] <- rst[[k]]@data@min
			Max[k] <- rst[[k]]@data@max
		}
		return(scale(rst, center=Min, scale=Max-Min))
	}
	# by default mean-std normalization
	return(scale(rst))
}


# Compute normalized difference (e.g. NDVI)
normdiff <- function(stk, red=3, nir=4, stack.it=FALSE, normalize=FALSE, norm="min-max") {
	require(raster)
	dims <- dim(stk)
	max.bands = max(red,nir)
	if (dims[3] < max.bands)
		stop("Raster file does not have enough bands\n")
	nirBand <- getValues(stk[[nir]])
	redBand <- getValues(stk[[red]])
	ndviBand <- (nirBand - redBand) / (nirBand + redBand)
	if (normalize == TRUE) {
		if (norm == "min-max") {
			center <- min(ndviBand)
			ampl <- max(ndviBand)-center
		} else {
			center <- mean(ndviBand)
			ampl <- sd(ndviBand)
		}
		ndviBand <- scale(ndviBand, center=center, scale=ampl)
	}
	ndvi <- raster(nrows=dims[1], ncols=dims[2], crs=crs(stk), ext=extent(stk), resolution=res(stk), vals=ndviBand)
	if (stack.it == FALSE)
		return(ndvi)
	return(stack(stk, ndvi))
}

# Moving function
# This function computes the moving (average, standard-deviation, etc) of a raster dataset
mov.fun <- function(rst, window.size, fun, normalize=FALSE, verbose=FALSE) {
	require(raster)
	if (!is.integer(window.size))
		window.size <- abs(ceiling(window.size))
	# need to be odd
	if (window.size%%2 == 0)
		window.size <- window.size + 1
	w <- matrix(1, nrow=window.size, ncol=window.size)
	n <- nlayers(rst)
	stk <- focal(rst[[1]], w, fun, na.rm=TRUE)
	if (verbose == TRUE)
		cat(">> Band ", 1," of ", n, " done.\n")
	for (k in 2:n) {
		tmp <- focal(rst[[k]], w, fun, na.rm=TRUE)
		stk <- stack(stk, tmp)
		if (verbose == TRUE)
			cat(">> Band ", k, " of ", n, " done.\n")
	}
	if (normalize == TRUE)
		stk <- raster.scale(stk)
	return(stk)
}

# function to calculate and plot kmeans for a raster object
performKMeans <- function(inputRaster, noClusters) {
	raster_df <- as.data.frame(inputRaster)
	clustering <- kmeans(raster_df, noClusters, iter.max = 100, nstart = 10)
	clusterRaster <- raster(inputRaster)
	clusterRaster <- setValues(clusterRaster, clustering$cluster)
	plot(clusterRaster)
	#return(clustering)
}

calculateError <- function(prediction, actual) {
	trainDiff <- predictionMatch - train[,7]
	trainDiffCount <- 0
	for(i in 1:length(trainDiff)) {
		if(trainDiff[i] != 0) {
			trainDiffCount <- trainDiffCount + 1
		}
	}
}
```

- Load a raster file, which has 4 layers
```{r}
rasterJ<-brick("C:\\Users\\Owner\\Documents\\Portugal\\Sat_Image_Classification\\data\\data\\img\\J_04AUG14112729-M2AS-000000137917_01_P001_etrs89.TIF")
#rasterJ<-brick("C:\\Users\\D059331\\Desktop\\DM GIC\\data\\img\\J_04AUG14112729-M2AS-000000137917_01_P001_etrs89.TIF")
rasterJ
```
- Plot map
```{r}
plotRGB(rasterJ, 3,2,1)
```

- choose a sector for clustering
```{r}
ext <- extent(-104294.4, -102964.5, -43623.48, -42742.44 )
sector <- crop(rasterJ, ext)
plotRGB(sector, 3, 2, 1)
```

- add normdiff layer for better vegetation recognition
```{r}
sector[[5]] <- normdiff(sector)
```
- plot each layer separately
```{r}
plot(sector)
```

- Get summary statistics of layers. Note that there are no NAs
```{r}
summary(sector)
```
- Input each of the five layers to the moving function. This creates windows of three by three pixels and calculates the mean and standard deviation of the pixels in these windows.Returns five rasters with these mean and standard deviation values filled in to the pixels.
```{r}
means<-mov.fun(sector,3,mean, normalize=FALSE, verbose=TRUE)
sds<-mov.fun(sector,3,sd, normalize=FALSE, verbose=TRUE)
```

- convert output of above to brick and summarize to check for na values
```{r}
stk<-stack(means,sds)
sectorMovings<-brick(stk)
summary(sectorMovings)
```

-We see that there are equal NA values in each layer
We get dimension information to determine number of cells in total in each layer
```{r}
sectorMovings
```

-We see that NA values are a very small percentage of total cells (less than one percent) of all cells#they are not visible in the plots
```{r}
sectorMovings
plot(sectorMovings)
```

-use trim to remove NA values
-from documentation: Trim (shrink) a Raster* object by removing outer rows and columns that all have the same value (e.g. NA). 
-Check output for NA values
```{r}

sectorMovingsTrim<-trim(sectorMovings, padding=0, values=NA)

```
-No NA values in any layers
-Plote output of trim to see if any issues
```{r}
sectorMovingsTrim<-trim(sectorMovings, padding=0, values=NA)
summary(sectorMovingsTrim)
plot(sectorMovingsTrim)
```
-It looks good. The moving function may have created NA values around edges, maybe where a full three by three grid could not be made.

-determine kmeans seed values
```{r}

```


-Normalize the ten layer stack because the range of values and variation differ between the RGB and infrared bands the NDVI. The five layers measuring the standard deviation also have a different range and variation level.
```{r}
sectorMovingsTrimNorm<-raster.scale(sectorMeansTrim)
sectorMovingsTrimNorm
```

-run kmeans on the normalized ten layer stack.
```{r}
performKMeans(sectorMovingsTrim,6)
```


-Assess the clusters we have found
-RAQUEL any writing about how we assess clusters (generally, or in R)
```{r}

```
-Add the the supervised report here
